---
title: "Logistic Regression Regularized"
---


```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(mlbench)
```

```{r sigmoid}
data <- mlbench.ringnorm(1000, d = 2)

data <- matrix(cbind(data$x[, 1], data$x[, 2], as.numeric(data$classes)),
               ncol = 3)
data[, 3] <- ifelse(data[, 3] == 1, 0, 1)
```
logistic regression hypothesis is defined as:

$h_\theta(x) = g(\theta^Tx)$

where function g is the sigmoid function. The sigmoid function is defined as:

$g(z) = \frac{1}{1+e^{-z}} = \frac{e^z}{1 + e^z}$

```{r sigmoid}
sigmoid <- function(z) {
    g <- 1 / (1 + exp(-z))
    return(g)
}

```

Logistic regression cost function:

$J(\theta) = \frac{1}{m} \displaystyle\sum_{i=1}^{m}[-y^{(i)} \log(h_\theta(x^{(i)})) - (1 - y^{(i)}) \log(1 - h_\theta(x^{(i)}))]$

```{r costFunction}
costFunction <- function(theta, X, y, lambda) {
    m <- length(y)
    J <- 0
    
    J <- 1/m * ((t(log(sigmoid(X %*% theta))) %*% (-y)) -
            t(log(1 - sigmoid(X %*% theta))) %*% (1 - y)) +
            lambda / (2 * m) * sum(theta[2:length(theta)] ^ 2)
    return(J)
}
```

```{r gradient}
gradient <- function(theta, X, y, lambda) {
    m <- length(y)
    grad <- rep(0, length(theta))
    
    grad <- 1/m * t(X) %*% (sigmoid(X %*% theta) - y) +
            rbind(0, matrix((lambda/m) * theta[2:length(theta)]))
    return(grad)
}
```

```{r}
data <- read.csv("C:\\Projects\\StanfordMachineLearning\\machine-learning-ex2\\ex2\\ex2data2.txt", header = FALSE)

colnames(data) <- c("Test1", "Test2", "QAResult")
X <- as.matrix(data)
y <- X[, 3]
X <- X[, 1:2]
##ggplot(data, aes(x = Test1, y = Test2, color = as.factor(QAResult), shape = as.factor(QAResult))) + geom_point(size = 2)
```

```{r}
ggplot(data, aes(x = Test1, y = Test2, color = as.factor(QAResult), shape = as.factor(QAResult))) + geom_point(size = 2)
```

```{r mapFeature}
mapFeature <- function(X1, X2) {
    degree <- 6
    out <- matrix(rep(1, length(X1)))
    for (i in 1:degree) {
        for (j in 0:i) {
            out <- cbind(out, (X1 ^ (i-j)) * (X2 ^ j))
        }
    }
    return(out)
}

```

```{r}
X <- mapFeature(X[, 1], X[, 2])
```

```{r}
initial_theta <- matrix(rep(0, ncol(X)))
lambda <- 100
theta_optim <- optim(par = initial_theta, fn = costFunction, gr = gradient, X = X, y = y, lambda = lambda, method = "BFGS")
```

This is the decision boundary of logistic regression that corresponds to the probability of 0.5
```{r decisionBoundary}
u <- v <- seq(-1, 1.5, length.out = 100)
z <- matrix(0, nrow =length(u), ncol = length(v))
for (i in 1:length(u)) {
    for (j in 1:length(v)) {
        z[i, j] <- mapFeature(u[i], v[j]) %*% theta_optim$par
    }
}

#ggplot(data, aes(x = Score1, y = Score2, color = as.factor(Admission), shape = as.factor(Admission))) + geom_point(size = 2) + geom_abline(slope = (-theta_optim$par[2] / theta_optim$par[3]), intercept = (-theta_optim$par[1] / theta_optim$par[3]), colour = "blue", size = 1)

# theta_optim$par
# https://stats.stackexchange.com/questions/6206/how-to-plot-decision-boundary-in-r-for-logistic-regression-model
# http://rstudio-pubs-static.s3.amazonaws.com/320280_2a9e44636b3a4f75a7dc02bf0b167572.html
```

Prediction:
For a student with scores 45 and 85, we predict an admission
#```{r prediction}
#sigmoid(c(1, 45, 85) %*% theta_optim$par)
#```

#```{r accuracy}
#prediction <- sigmoid(X %*% theta_optim$par) >= 0.5
#mean(as.integer(prediction) == y) * 100
#```


```{r}
rownames(z) <- seq(from = -1, to = 1.5, length.out = 100)
colnames(z) <- seq(from = -1, to = 1.5, length.out = 100)

as.data.frame(z) %>% 
  rownames_to_column() %>% 
  gather(key, value, -rowname) %>% 
  mutate(key = as.numeric(key), 
         rowname = as.numeric(rowname)) %>%
  ggplot() + geom_contour(aes(x = rowname, y = key, z = value), bins = 1) + 
    geom_point(data=data, aes(x = Test1, y = Test2, color = as.factor(QAResult),
                              shape = as.factor(QAResult)), size = 2)
```